---
title: "Exponential distribution comparison with Central Limit Theorem"
author: "Hector Mario Romer"
date: "2/12/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 3, 
                      message = FALSE, warning = FALSE)
```
# Statistical Inference Course Project pt 1
## Overview

In this project I am going to investigate the exponential distribution in R an compare it with the CLT (Central Limit Theorem). Lambda will be set to 0.2 for all of the simulations. I will investigate the distribution of averages of 40 exponentials, doing a thousand simulations. 

## Simulations

Run the test and save the result in a matrix of size n * tests called *exp_dist*. The means are saved in dataframe *exp_dist_means*.

```{r}
# Load required libraries
suppressMessages(library(ggplot2))

# Set seed 
set.seed(1307)

# Set constants as required
lambda <- 0.2
n <- 40
tests <- 1000

# run the test
exp_dist <- matrix(data = rexp(n * tests, lambda), nrow = tests)
exp_dist_means <- data.frame(means = apply(exp_dist, 1, mean))
```

## Sample Mean versus Theoretical Mean

Knowing that the expected mean $\mu$ of a exponential distribution of rate $\lambda$ is: 


$\mu= \frac{1}{\lambda}$ 


Calculate the expected mean *mu*:

```{r}
mu <- 1 / lambda
mu
```


Calculate $\bar X$ as the average sample mean of 1000 simulations of 40 randomly sampled exponential distributions:

```{r}
X <- mean(exp_dist_means$means)
X
```


We can see that the average sample mean (```r X```) is very close to the expected mean (```r mu```)


Plotting the means:

```{r}
ggplot(data = exp_dist_means, aes(x = means)) +
        geom_histogram(binwidth = 0.1, color = "navy", fill = "lightblue") +
        theme_bw() +
        scale_x_continuous(breaks = round(seq(min(exp_dist_means$means), 
                max(exp_dist_means$means), by = 1)))
```

## Sample Variance versus Theoretical Variance

The expected standard deviation $\sigma$ of a exponential distribution rate $\lambda$ is:


$\sigma = \frac{1/\lambda}{\sqrt{n}}$ 


Therefore, our expected standard deviation *sd* is:

```{r}
sd <- (1/lambda) / sqrt(n)
sd
```


The variance $Var$ of standard deviation $\sigma$ is:

$Var = \sigma^2$

```{r}
Var <- sd^2
Var
```


Having $Var_x$ as the variance of the average sample mean of 1000 simulations of 40 randomly sampled exponential distribution, and $\sigma_x$ as the corresponding standard deviation:

```{r}
Var_x <- var(exp_dist_means$means)
Var_x

sd_x <- sd(exp_dist_means$means)
sd_x
```


We can see that both the variance are the standard deviaton are close to the expected values. Please note that the variance is the sd squared, which makes minor differences bigger.

## Distribution

Let's compare the population means and standard deviation (in blue) with a normal distribution of the expected values (in red). Vertical lines represent means: expected (red) and calculated (blue).

```{r}
ggplot(data = exp_dist_means, aes(x = means)) + 
        theme_bw() +
        geom_histogram(binwidth = 0.1, aes(y =..density..), alpha = 0.2) + 
        stat_function(fun = c("dnorm"), args = list(mean = mu , sd = sd),
              colour = "tomato3", size = 1) + 
        geom_vline(xintercept = mu, size = 1, colour = "tomato3") + 
        geom_density(colour = "navy", size = 1) +
        geom_vline(xintercept = X, size = 1, colour="navy") + 
        scale_x_continuous(breaks = seq(mu-3, mu+3, 1), limits = c(mu-3,mu+3)) 
```

We can see that the calculated distributions of means of random sampled exponential distributions are close to the expected values using the given lambda.