---
title: "ToothGrowth: Basic inferential data analysis"
author: "Hector Mario Romer"
date: "2/13/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
# Statistical Inference Course Project pt. 2

## Overview

In this part of the Project, I am going to load the ToothGrowth data and perform some basic exploratory data analyses. A basic summary of the data will be provided. Confidence intervals and/or hypothesis will be use to compare tooth growth by supp and dose. Conclusions will be stated based on assumptions draw from the analyses.


## Load Data

Load the dataset ToothGrowth and stores it in *tg*. Print the structure of the dataset, followed by the first five rows. Variable "len" is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).

```{r}
# load required libraries
library(ggplot2)
library(datasets)
suppressMessages(library(dplyr))

data(ToothGrowth)
tg <- ToothGrowth
str(tg)
head(tg, 5)
```


## Basic Summary of the data

Using the summary function to summarize the dataset:

```{r}
summary(tg)
```

We can see that 30 observations correspond to orange juice (supp = "OJ") and 30 observations to vitamin C (supp = "VC"). Since "dose" levels are fixed quantities, its mean and median carry little information. 

Using a boxplot to visualize the impact (if any) of dossage and supplement on tooth growth:

```{r}
library(ggplot2)
p <- ggplot(tg, aes(x = dose, y = len)) + 
        theme_bw() +
        geom_boxplot(aes(fill = factor(dose))) +
        geom_point() +
        facet_grid(.~supp) +
        ggtitle("Tooth growth by supplement / dosage")
p
```

Looking at the chart, it appears that when dose is 2mg the median for OJ and VC are similar. When the dose is lower (0.5mg and 1mg) orange juice has a much possitive impact on tooth growth.

## Hypothesis tests

Let's consider the following two hypothesis:

H0 = Orange juice has an impact of tooth growth compared to ascorbic acid

H1 = Orange juice does not have an impact of tooth growth compared to ascorbic acid

```{r}
t.test(len ~ supp, data = tg)
t.test(len ~ supp, data = tg[tg$dose < 2,])
t.test(len ~ supp, data = tg[tg$dose == 2,])
```

The p-value on the first test is 0.06, so since the value is greater than 0.05 and the confidence interval of the test contains zero, we can say that orange juice has no significant impact in tooth growth compared to ascorbic acid.

In the second test, p-value is 0.004 and the confidence interval does not contains zero. Therefore, we can say that when dosage is below 2mg, orange juice has a greater impact on tooth growth compared to ascorbic acid.

Third test, has p-value equal to 0.96 and zero almost in the middle of the confidence interval. This tells us that at 2mg dosage, there is no significant difference between orange juice and ascorbic acid.

## Assumption and Conclusions

Assuming that the supplements were distributed identically and independently among the guinea pigs in the study, we conclude that orange juice in dosages below 2 mg have a greater impact on tooth growth of guinea pigs compared to ascorbic acid. When dosage quantity is 2mg, there is no significant difference.
